---
title: "R Notebook"
output: html_notebook
---
```{r}
library(ggplot2) #load the library
data(msleep) #load a data set on mamallian sleep patterns
str(msleep)
```

Asthetics *aes()* contorl how data is mapped onto plot
  eg: aes(x=age, y=height, shape=gender)
Geometries *geom_* specify type of plot
  eg: geom_points()
  
Want a **scatterplot** b/t sleep and brainweight
```{r}
pl<- ggplot(data = msleep, mapping= aes(x=brainwt, y=sleep_total))
pl<- pl +geom_point()
pl
```

```{r}
pl<- ggplot(data = msleep, mapping= aes(x=brainwt, y=sleep_total))
pl<- pl +geom_point()
pl
pl+scale_x_log10()
```
*pl + scale* just shows the plot plus that modification
To permanently modify object, have to assign modification.

```{r}
pl<- ggplot(data = msleep, mapping= aes(x=brainwt, y=sleep_total))
pl<- pl +geom_point()
pl<-pl+ scale_x_log10()
pl
```

```{r}
pl<- ggplot(data = msleep, mapping= aes(x=brainwt, y=sleep_total))
pl<- pl +geom_point()
pl<-pl+ scale_x_log10()
pl<- pl+ aes(color=vore)
pl
```

This colors the points by another column in the data, such as "vore" in this example.

```{r}
pl2<-ggplot(data=msleep, mapping=aes(x=bodywt, y=sleep_total))
pl2<-pl2+geom_point()
pl2<-pl2+scale_x_log10()
pl2
```

**BoxPlots**
```{r}
bl<-ggplot(data=msleep, mapping=aes(x=vore, y=sleep_rem))
bl<- bl+geom_boxplot()
bl
```


```{r}
bl<-ggplot(data=msleep, mapping=aes(x=vore, y=sleep_rem))
bl<- bl+geom_boxplot()
bl2<- bl +geom_point(position=position_jitter(width=.3))
bl2
```