---
title: "Intro to R Notebook"
output: html_notebook
---

## A silly R program

This is the first program that Professor Maloof asked us to write in R

'''
{r hello_world2}
print("Hello World")
'''

## swirl tutorial

> cbind(patients, my_matrix)
     patients                       
[1,] "Bill"   "1" "5" "9"  "13" "17"
[2,] "Gina"   "2" "6" "10" "14" "18"
[3,] "Kelly"  "3" "7" "11" "15" "19"
[4,] "Sean"   "4" "8" "12" "16" "20"

| Your dedication is inspiring!

  |=====================================                        |  61%

| Something is fishy about our result! It appears that combining the
| character vector with our matrix of numbers caused everything to be
| enclosed in double quotes. This means we're left with a matrix of
| character strings, which is no good.

...

  |=======================================                      |  64%

| If you remember back to the beginning of this lesson, I told you
| that matrices can only contain ONE class of data. Therefore, when we
| tried to combine a character vector with a numeric matrix, R was
| forced to 'coerce' the numbers to characters, hence the double
| quotes.

...

  |=========================================                    |  67%

| This is called 'implicit coercion', because we didn't ask for it. It
| just happened. But why didn't R just convert the names of our
| patients to numbers? I'll let you ponder that question on your own.

...

  |==========================================                   |  69%

| So, we're still left with the question of how to include the names
| of our patients in the table without destroying the integrity of our
| numeric data. Try the following -- my_data <- data.frame(patients,
| my_matrix)

> my_data<-data.fram(patients, my_matrix)
Error: could not find function "data.fram"
> ## Exercise 6
_When extracting from a 2-dimensional object, which number specifies rows and which specifies columns?_

```{r}
m<-matrix(data=1:25,ncol=5,byrow=T)
m
m[5,5]
```

     [,1] [,2] [,3] [,4] [,5]
[1,]    1    2    3    4    5
[2,]    6    7    8    9   10
[3,]   11   12   13   14   15
[4,]   16   17   18   19   20
[5,]   21   22   23   24   25
[1] 25

ncol is the column
nrow is for the rows
[row, column]

_What does `m[3,]` do?_

```{r}
m[3,]
```

[1] 11 12 13 14 15
It goes to the 3rd row of your matrix and displays it.

_How can you extract the 3rd, 4th and 5th columns of m together as one object?_

```{r}
m[,3:5]
```

    [,1] [,2] [,3]
[1,]    3    4    5
[2,]    8    9   10
[3,]   13   14   15
[4,]   18   19   20
[5,]   23   24   25

## Exercise 7
_What does the cbind command do?  How about rbind?_

```{r}
cbind(m,101:105)
```

     [,1] [,2] [,3] [,4] [,5] [,6]
[1,]    1    2    3    4    5  101
[2,]    6    7    8    9   10  102
[3,]   11   12   13   14   15  103
[4,]   16   17   18   19   20  104
[5,]   21   22   23   24   25  105

cbind combines the vectors and matrices by their columns (or adds a column of numbers you specify).
rbind combines them by rows.

_Create a new object "n" where the first row is a new row of numbers (your choice) and the following rows are the followed by the m matrix._

```{r}
n<-c(4,6,7,12,98)
o<-rbind(n,m)
o
o[2,]
```

  [,1] [,2] [,3] [,4] [,5]
n    4    6    7   12   98
     1    2    3    4    5
     6    7    8    9   10
    11   12   13   14   15
    16   17   18   19   20
    21   22   23   24   25
[1] 1 2 3 4 5

Note: The matrix did not have names, so when the row "n" was inserted, the index for rows disappeared. It still recognizes the row order.


| Excellent work!

  |============================================                 |  72%

| Now view the contents of my_data to see what we've come up with.

> my_data
  patients X1 X2 X3 X4 X5
1     Bill  1  5  9 13 17
2     Gina  2  6 10 14 18
3    Kelly  3  7 11 15 19
4     Sean  4  8 12 16 20

| That's a job well done!

  |==============================================               |  75%

| It looks like the data.frame() function allowed us to store our
| character vector of names right alongside our matrix of numbers.
| That's exactly what we were hoping for!

...

  |===============================================              |  78%

| Behind the scenes, the data.frame() function takes any number of
| arguments and returns a single object of class `data.frame` that is
| composed of the original objects.

...

  |=================================================            |  81%

| Let's confirm this by calling the class() function on our newly
| created data frame.

> class(my_data)
[1] "data.frame"

| Keep up the great work!

  |===================================================          |  83%

| It's also possible to assign names to the individual rows and
| columns of a data frame, which presents another possible way of
| determining which row of values in our table belongs to each
| patient.

...

  |=====================================================        |  86%

| However, since we've already solved that problem, let's solve a
| different problem by assigning names to the columns of our data
| frame so that we know what type of measurement each column
| represents.

...

  |======================================================       |  89%

| Since we have six columns (including patient names), we'll need to
| first create a vector containing one element for each column. Create
| a character vector called cnames that contains the following values
| (in order) -- "patient", "age", "weight", "bp", "rating", "test".

> c("patient","age","weight","bp","rating","test")
[1] "patient" "age"     "weight"  "bp"      "rating"  "test"   

| Not quite, but you're learning! Try again. Or, type info() for more
| options.

| Make sure all of the names are lowercase, surrounded by double
| quotes, and separated with commas. Don't forget to use the c()
| function.

> cnames<-c("patient","age","weight","bp","rating","test")

| You are quite good my friend!

  |========================================================     |  92%

| Now, use the colnames() function to set the `colnames` attribute for
| our data frame. This is similar to the way we used the dim()
| function earlier in this lesson.

> colnames(cnames)
NULL

| Not quite! Try again. Or, type info() for more options.

| Try colnames(my_data) <- cnames.

> colnames(my_data)<-cnames

| Perseverance, that's the answer.

  |==========================================================   |  94%

| Let's see if that got the job done. Print the contents of my_data.

> my_data
  patient age weight bp rating test
1    Bill   1      5  9     13   17
2    Gina   2      6 10     14   18
3   Kelly   3      7 11     15   19
4    Sean   4      8 12     16   20

| You nailed it! Good job!

  |===========================================================  |  97%

| In this lesson, you learned the basics of working with two very
| important and common data structures -- matrices and data frames.
| There's much more to learn and we'll be covering more advanced
| topics, particularly with respect to data frames, in future lessons.

class() to see if data frame or matrix or vector
dim()to see exactly how many rows and columns we're dealing with.

names() will return a character vector of column (i.e. variable) names.
summary() provides different output for each variable, depending on its class. For numeric data such as Precip_Min, summary() displays the minimum, 1st quartile, median, mean, 3rd quartile, and maximum. These values help us understand how the data are distributed.
 For categorical variables (called 'factor' variables in R), summary() displays the number of times each value (or 'level') occurs in the data. For example, each value of Scientific_Name only appears once, since it is unique to a specific plant. In contrast, the summary for Duration (also a factor variable) tells us that our dataset contains 3031 Perennial plants, 682 Annual plants, etc.
 

str() combines many of the features of the other functions you've already seen, all in a concise and readable format. At the very top, it tells us that the class of plants is 'data.frame' and that it has 5166 observations and 10 variables. It then gives us the name and class of each variable, as well as a preview of its contents.

----------------------------

## Exercise 6
_When extracting from a 2-dimensional object, which number specifies rows and which specifies columns?_

```{r}
m<-matrix(data=1:25,ncol=5,byrow=T)
m
m[5,5]
```

     [,1] [,2] [,3] [,4] [,5]
[1,]    1    2    3    4    5
[2,]    6    7    8    9   10
[3,]   11   12   13   14   15
[4,]   16   17   18   19   20
[5,]   21   22   23   24   25
[1] 25

ncol is the column
nrow is for the rows
[row, column]

_What does `m[3,]` do?_

```{r}
m[3,]
```

[1] 11 12 13 14 15
It goes to the 3rd row of your matrix and displays it.

_How can you extract the 3rd, 4th and 5th columns of m together as one object?_

```{r}
m[,3:5]
```

    [,1] [,2] [,3]
[1,]    3    4    5
[2,]    8    9   10
[3,]   13   14   15
[4,]   18   19   20
[5,]   23   24   25

## Exercise 7
_What does the cbind command do?  How about rbind?_

```{r}
cbind(m,101:105)
```

     [,1] [,2] [,3] [,4] [,5] [,6]
[1,]    1    2    3    4    5  101
[2,]    6    7    8    9   10  102
[3,]   11   12   13   14   15  103
[4,]   16   17   18   19   20  104
[5,]   21   22   23   24   25  105

cbind combines the vectors and matrices by their columns (or adds a column of numbers you specify).
rbind combines them by rows.

_Create a new object "n" where the first row is a new row of numbers (your choice) and the following rows are the followed by the m matrix._

```{r}
n<-c(4,6,7,12,98)
o<-rbind(n,m)
o
o[2,]
```

  [,1] [,2] [,3] [,4] [,5]
n    4    6    7   12   98
     1    2    3    4    5
     6    7    8    9   10
    11   12   13   14   15
    16   17   18   19   20
    21   22   23   24   25
[1] 1 2 3 4 5

Note: The matrix did not have names, so when the row "n" was inserted, the index for rows disappeared. It still recognizes the row order.

qb: where align start
qe: qhere align end

_Want more? (OPTIONAL) do the same but reverse the order of the rows from m._



### Ciera's Tutorial
Activity Break
There are many ways to subset data and you can get a brief look with clear examples here.

1. Subset the data to only nerds and metalheads.

```{r}
nerdsandmetalheads<-subset(stereotypes, population=="nerd" | population=="metalheads")
```

2. Use what you have learned and this page as guidance for sub-setting a dataset to answer the question: Are there more males who binge drink (over 25 a week) and eat over 20 tacos a week in the metal heads or hipster population?

```{r}
hipsters<-subset(stereotypes, population=="hipster")
hipstermales<-subset(hipsters, gender=="male")
hipsterbingedrink<-subset(hipstermales, beer > 25)
hipsterbingeeats<-subset(hipsterbingedrink, tacos >20)
```

```{r}
metalheads<-subset(stereotypes, population=="metalhead")
mhmales<-subset(metalheads, gender=="male")
mhbingedrink<-subset(mhmales, beer > 25)
mhbingeeats<-subset(mhbingedrink, tacos >20)
```

3. Ask one more question about the dataset that you are interested in. Subset the data accordingly.

*Do females or males shower more often on average?*

```{r}
fem<-subset(stereotypes, gender == "female")
mal<-subset(stereotypes, gender=="male")
femshower<-sort(fem$shower)
malshower<-sort(mal$shower)
mean(fem$shower)
mean(mal$shower)
```


4. In your own words explain what “==” means?

Find the word or value in the file.

# Visualization

__Visualization with ggplot2___

View the distribution of showers. Color by population.
```{r}
qplot(shower, data=stereotypes, geom="density",alpha = I(0.5), fill=population)
```

View the distribution of beer consumption and color by gender.
```{r}
qplot(beer, data=stereotypes, geom="density",alpha = I(0.5), fill=gender)
```

Find a way to create a plot using the the geom violin with the stereotype data set.
```{r}
s<-ggplot(data=stereotypes, aes(x=population, y=shower)) + geom_violin()
s
```

__Combining Subset with Visualization__

1.Make a graph that visually describes the difference between beer consumption between gender in the nerd category.

```{r}
nerds<- subset(stereotypes, population == "nerd")
ggplot(data=nerds, aes( x=gender, y=beer, colour=gender)) +geom_boxplot()
ggplot(data=stereotypes, aes(x=gender, y=beer)) +geom_boxplot()
```
The last command is the control/comparison group that has all the populations.


2.Using theggplot2 documentation, explore three different geoms, using either the full data set or a subset of the data. Tell me something interesting about the data.

```{r}
ggplot(stereotypes, aes(beer, tacos, color = population, shape = gender)) + geom_point()
males<- subset(stereotypes, gender == "male")
females<- subset(stereotypes, gender == "female")
ggplot(data=males, aes( x=population, y=computer, colour=population)) +geom_boxplot()
ggplot(data=females, aes( x=population, y=computer, colour=population)) +geom_boxplot()

ggplot(data=stereotypes, aes( x=gender, y=computer, colour=population)) +geom_boxplot()
```
All males, except male nerds, on average spend more time than their female population counterparts on the computer.

3.From your exploration of the stereotypes data make three hypotheses about the data.

Nerds stay indoors more often on the computer and so they have more immediate access to the shower.

Tacos and Beer are meant to be together (they're offered at the same places).

Computers are usually marketed as to males, so they usually spend more time on the computers.

__Visualization with the ggplot() function__

With the p base layer, make a unique plot by adding additional layers.
```{r}
p <- ggplot(stereotypes, aes(shower, fill = gender))
p + geom_bar()
p + geom_histogram(binwidth = 1, position = "dodge")
p + geom_histogram(binwidth = 1, position = "dodge") + coord_flip()
p + geom_histogram(binwidth = 5, position = "dodge") + facet_grid(~population)

p <- ggplot(stereotypes, aes(tacos, fill = gender))
p+ geom_histogram(binwidth=5,position = "dodge") +facet_grid(~age)
```
Make new base layer. Explore what the difference is between color = and fill =. Use your own words to discern the difference.

*fill* fills the whole bar with color
*colour* just assigns color to the variables but it doesn't necessarily completely fill it in.

__Visualization with the ggplot() function with more color adjustments__

Read about and use the following modifiers correctly with the any portion of the sterotypes dataset you want.

1.Modify color with scale_color_hue()

```{r}
ggplot(stereotypes, aes(age, tacos, color = population)) + 
    geom_point()+ scale_color_hue()
```

2.Make a tricked out box plot using the stereotypes data. Use at least 3 of the aesthetic options available for geom_boxplot. Use ggplot2 boxplot documentation for guidance.

(used facid grid, outlier coloring, scale_fill_hue, and adding the notch)

```{r}
ggplot(stereotypes, aes(age, tacos, color = population)) + 
    geom_boxplot(outlier.color = "red",notch=TRUE)+scale_fill_hue()+facet_grid(~population)
```

## Exercise 12
_Use ggplot to explore the relationship between score ("Score") and alignment length("len").  Provide a plot that illustrates the relationship and describe the relationship in words._
```{r}
fly.worm.small<-fly.worm[sample(nrow(fly.worm), 1000),]
ggplot(fly.worm.small, aes(len,Score))+geom_point()
```
Roughly, it looks like alignment length and score are positively correlated, but the variance seems fairly large. The variance is especially strong with lengths under 2000.

## Excercise 13
_While you might expect that BLAST results with long alignments would have high scores, this is not always the case.  Form a hypothesis as to what might influence the relationship between alignment length and score.  Use ggplot to make a new plot to explore this hypothesis.  Does the plot support your hypothesis or not?  State your hypothesis, provide the code for your plot, and state your conclusion._
```{r}
ggplot(fly.worm.small, aes(len, Score, color=gaps)) + geom_point()
```

Based on the graph, there is a small group between 1000-2000 length with a lot of gaps that also have low scores, supporting my hypothesis. However, despite varying gap amounts, relatively long lengths(over 2000) still have relatively high scores. Low length alignments with low amount of gaps can range from low scoring to high scoring, so there must be something else affecting this relationship. My hypothesis does not completely explain the pattern.
------------------------

Lines that begin with a pound sign # are comment lines and are ignored by R.
